---
title: "Homework 3 - Visualization and EDA"
author: "Kristi Chau"
date: "10/5/2019"
output: github_document
fig_caption: yes
latex_engine: xelatex
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage[normalem]{ulem}
   - \usepackage[utf8]{inputenc}
   - \usepackage{makecell}
   - \usepackage{xcolor}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(knitr)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1

```{r, include = FALSE}
library(p8105.datasets)
data("instacart")
```

The `instacart` data set has **`r nrow(instacart)`** observations and **`r ncol(instacart)`** variables, which gives **`r nrow(instacart) * ncol(instacart)`** distinct datapoints. There are **`r sum(is.na(instacart))`** missing datapoints in this dataset. Some key variables of this dataset are: **`r colnames(instacart)[1]`**, which ensures all items ordered are associated with the same order, **`r colnames(instacart)[5]`**, who placed the order, **`r colnames(instacart)[11]`**, what product was ordered, and **`r colnames(instacart)[15]`**, what department the product is from. The most orders made by a customer is **`r max(instacart[7])`**. The most popular item ordered is **`r names(which.max(table(instacart[11])))`**.

*1a: How many aisles are there, and which aisles are the most items ordered from?*

```{r}

count_aisles = instacart %>%
  distinct(aisle) %>% 
  count()

most_items_aisle = instacart %>% 
  count(aisle) %>%
  slice(which.max(n))
```

**There are `r pull(count_aisles,n)` aisles total. The aisle with the most items ordered is `r pull(most_items_aisle,aisle)` with a total of `r pull(most_items_aisle,n)` items ordered from that aisle.**

*1b: Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.*

```{r}
instacart_plot = instacart %>% 
  group_by(aisle) %>% 
  mutate(freq = n()) %>% 
  ungroup() %>% 
  mutate(
    aisle = as.factor(aisle),
    aisle = fct_reorder(aisle, freq, na.rm = TRUE)
  )  %>%
  filter(freq > 10000) %>% 
  select(aisle,freq,department) %>% 
  distinct() %>% 
  ggplot(aes(x = aisle, y = freq/1000, fill = department)) +
  geom_col(position = "dodge") +
  labs(
    title = "Number of Items Ordered, By Aisle",
    y = "Number of Items Ordered (in 1000s)",
    x = "Aisle Name",
    caption = "Data from Instacart") +
  theme(
    axis.text.x = element_text(angle = 90,
                               hjust = 1),
    legend.position = "right") + 
  scale_fill_discrete(name = "Department Name") +
  scale_y_continuous(
  breaks = c(10,25,50,75,100,150), 
  labels = c("10","25","50","75","100","150"))

instacart_plot
```

**10,000+ items were ordered from 39 aisles out of the total 134 aisles. Out of these 39 aisles, the `butter` aisle had the least items ordered, while `fresh fruits` and `fresh vegetables` had over 150,000 items ordered, far surpassing the 3rd ranked aisle, which was `packaged vegetables fruits`, with 78,493 items ordered.**

*1c: Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.*

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle,product_name) %>%
  mutate(freq = n()) %>%
  arrange(aisle,desc(freq)) %>%
  distinct(aisle,product_name,freq) %>%
  mutate(prod_freq = paste(product_name,freq, sep = "\n")) %>%
  ungroup() %>%
  group_by(aisle) %>%
  mutate(rank = dense_rank(desc(freq))) %>%
  filter(rank %in% c(1,2,3)) %>%
  select(rank,aisle,prod_freq) %>% 
  pivot_wider(
    names_from = "aisle",
    values_from = "prod_freq"
  ) %>% 
  kable(
    caption = "Most Popular Items and Number of Times Each Item was Ordered in 'Baking Ingredients', 'Dog Food Care', and 'Packaged Vegetables Fruits' Aisles",
    col.names = c("Ranking",
                  "Baking Ingredients",
                  "Dog Food Care",
                  "Packaged Vegetables Fruits")
  )
```

**The most popular `Baking Ingredients` items (and frequency of purchase) are `Light Brown Sugar (499)`, `Pure Baking Soda (387)`, and `Cane Sugar (336)`. The most popular `Dog Food Care` items (and frequency of purchase) are `Snack Sticks Chicken & Rice Recipe Dog Treats (30)` `Organix Chicken & Brown Rice Recipe (28)`, and `Small Dog Biscuits (26)`. The most popular `Packaged Vegetables Fruits` items (and frequency of purchase) are `Organic Baby Spinach (9784)`, `Organic Raspberries (5546)`, and `Organic Blueberries (4966)`.**

*1d: Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).*

```{r}
instacart %>%
  select(product_name,order_dow,order_hour_of_day) %>%
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>%
  group_by(product_name,order_dow) %>%
  mutate(mean_hod = mean(order_hour_of_day)) %>% 
  distinct(product_name,order_dow,mean_hod) %>%
  arrange(order_dow) %>%
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hod"
  ) %>%
  kable(
    caption = "Mean hour of the day, by day of the week, at which Pink Lady Apples and Coffee Ice Cream are ordered",
    col.names = c("Product Name",
                  "Sunday",
                  "Monday",
                  "Tuesday",
                  "Wednesday",
                  "Thursday",
                  "Friday",
                  "Saturday")
  )
```

**`Pink Lady Apples`, on average, are ordered the earliest on Mondays at `11.36` hours and the latest on Wednesdays at `14.25` hours. `Coffee Ice Cream`, on average, is ordered the earliest on Fridays at `12.26` hours and the latest on Tuesdays at `15.38` hours.**



# Problem 2

```{r, include = FALSE}
library(p8105.datasets)
data("brfss_smart2010")
```

*Data Cleaning*

```{r}
brfss_df = 
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  arrange(factor(response, levels =  c("Poor","Fair","Good","Very good","Excellent")))
```

**I filtered the data to contain only the Overall Health topic and arranged the data so responses were ordered from "Poor" to "Excellent" for the question "How is your general health?" There are `r nrow(brfss_df)` observations in the dataset, with `r ncol(brfss_df)` variables. There are `r sum(is.na(brfss_df))` missing datapoints in this dataset. The average sample size over all locations, across all states, was `r mean(pull(brfss_df,sample_size))`. This data was collected from 2002 to 2010.**

*2a: In 2002, which states were observed at 7 or more locations? What about in 2010?*

```{r, results = FALSE}
brfss_df %>%
  filter(year == 2002) %>%
  distinct(locationabbr,locationdesc) %>%
  count(locationabbr) %>%
  filter(n >= 7)
```

**In 2002, `Connecticut, Florida, Massachusetts, North Carolina, New Jersey, and Pennsylvania` were observed at 7 or more locations.**

```{r, results = FALSE}
brfss_df %>%
  filter(year == 2010) %>%
  distinct(locationabbr,locationdesc) %>%
  count(locationabbr) %>%
  filter(n >= 7)
```

**In 2010, `California, Colorado, Flordia, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas, and Washington` were observed at 7 or more locations.**

*2b: Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).*

```{r}
brfss_df %>%
  filter(response == "Excellent") %>%
  select(year,locationabbr,data_value) %>%
  group_by(year,locationabbr) %>%
  mutate(avg_data_value = mean(data_value)) %>%
  ggplot(aes(x = year, y = avg_data_value, color = locationabbr)) +
  geom_line(size = 1.25, alpha = 0.8) +
  labs(
    title = "Mean Prevalence of Landlines, Among Those who Responded 'Excellent', Over Time, by State",
    x = "Year",
    y = "Average Prevalence of Landlines",
    caption = "BRFSS data") +
  theme(
   legend.position = "right") +
  scale_colour_discrete (name = "State (and DC)")
```

**Given that we are observing the mean data value across time for all states in this dataset, it is difficult to see if there are any trends of prevalence of landline telephones over time among those who responded "Excellent" unless we focus on individual states or on a smaller subset of states, like region.**

*2c: Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.*

```{r}
brfss_df %>%
  filter(year %in% c(2006, 2010), locationabbr == "NY") %>%
  ggplot((aes(x = reorder(response,desc(display_order)), y = data_value, fill = locationdesc))) +
  geom_col(position = "dodge") + 
  labs(
    title = "Prevalence of Landlines in NY State in 2006 and 2010, by Location",
    x = "Response to 'How is your general health?'",
    y = "Prevalence of Landlines",
    caption = "BRFSS data") +
  facet_grid(~year) +
  theme(legend.position = "right") +
  scale_y_continuous(
    breaks = c(10,15,20,25,30,35,40,45))
```

**The distribution of data_value (the prevalence of landlines) for responses among locations in NY State seem to follow similar patterns between 2006 and 2010. The prevalence of landlines was lowest among those who responded that their general health was "poor"Poor" across all locations in 2006 and 2010, followed by "Fair" and "Excellent" responses. Depending on location, the prevalence of landlines was highest in those who responded "Good" or "Very good".**



# Problem 3

*3a: Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).*

```{r}
accel_data = read_csv("./data/accel_data.csv", col_types = cols(
  week = col_integer(),
  day_id = col_integer()
)) %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    names_prefix = "activity_",
    names_ptypes = list(activity = integer()),
    values_to = "counts"
  ) %>% 
  mutate(day_end = ifelse(day %in% c("Saturday","Sunday"),"weekend","weekday"))
```

**The variables in this dataset are `r colnames(accel_data)[1:6]` There are `r nrow(accel_data)` rows and `r ncol(accel_data)` columns for a total of `r nrow(accel_data) * ncol(accel_data)` observations. This patient was observed for `r max(pull(accel_data,day_id))` days and had a mean activity count of `r round(mean(pull(accel_data,counts)),2)` per minute.**

*3b: Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?*

```{r}
accel_data %>%
  mutate(
    day = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", ordered = TRUE)),
    day_num = as.integer(day)
  ) %>%
  group_by(week, day_num) %>%
  mutate(total_activity = sum(counts)) %>%
  ungroup() %>%
  arrange(week,day_num) %>%
  distinct(week,day,total_activity) %>%
  pivot_wider(
    names_from = "day",
    values_from = "total_activity"
  ) %>% 
  kable(
    caption = "Total Daily Activity over 5 weeks"
  )
```

**There are no obvious trends in this data. Tuesdays, Wednesdays, and Thursdays, the patient seems to have a somewhat consistent level of activity, while activity level varies more widely for Sunday, Monday, Friday, and Saturday. On the last 2 Saturdays of observation, it appears that the patient was not active.**

*3c: Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.*

```{r, message = FALSE}
accel_data %>%
  mutate(
    day = ordered(day,levels = c("Sunday", "Monday", 
    "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))) %>%
  ggplot(aes(x = activity, y = counts, color = day)) +
  geom_smooth(size = 2) +
  labs(
    title = "Average 24-hour Activity Course, by Day",
    x = "Hour",
    y = "Average Hourly Activity",
    caption = "Data from CUMC") +
  theme(
    legend.position = "right",
    panel.grid.major.x = element_line(color = "grey50")
  ) + 
  scale_colour_discrete(name = "Day of the Week") +
  scale_x_continuous(
  breaks = c(60,120,180,240,300,360,420,480,540,600,660,720,780,840,900,960,1020,1080,1140,1200,1260,1320,1380,1440), 
  labels = c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23","24"))
```

**On average, there was a period of high activity on Friday nights, peaking around 9-9:30pm. On Sundays, there was a period of high activity that peaked around 10:30-11am. On average, Saturdays appear to have the least amount of activity across the 5 weeks.**

